Literature Report on DP-means

### 2015-06-10 stat302/paper_report.tex
Done first draft of paper report

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

@@

\newcommand{\st}{ \; \big | \:}

@@

\begin{document}
\title{Literature Report on DP-means}
\author{Aaron Maurer}
\date{STAT 302, Spring Quarter 2015}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
For this report, I chose to read "Revisiting k-means: New Algorithms via Bayesian Nonparametrics"\footnote{Brian Kulis and Michael I. Jordan, ``Revisiting k-means: New Algorithms via Bayesian Nonparametrics'', \textit{CoRR} (2011): http://arxiv.org/abs/1111.0352}. This article introduces a alternative to the k-means clustering algorithm called DP-means, where instead of specifying a number of clusters at the start, then working to optimize their fit, new clusters are introduced when a point in the data set is too far from one of the existing clusters. This in itself is a `hard' clustering algorithm, which outputs an assignment to clusters, rather than a posterior distribution one might normally expect from a Bayesian method. However, this paper demonstrates that this algorithm is the limit of a Bayesian model where the total distribution is a mixture of Gaussian distributions, with the number of components coming from a Dirichlet process. This is the source of the name of the algorithm: Dirichlet process means or DP-means. \par
In addition to a description and derivation of this algorithm, the paper offers a few extensions and simulations. The most interesting of theses is a hierarchical clustering method where, over multiple data sets of the same variables, local clusters are simultaneously fit on each data set so as to match a set of global clusters across all of the data sets. This is again a hard clustering algorithm, but it can be derived as well in a Bayesian fashion by taking the limit when the parameters of the Dirichlet process which generates the components in each data set arises from a prior global Dirichlet process. The next extension shows how DP-means extends to spectral clustering; where one would spectrally cluster using k-means by performing k-means on the first k eigenvectors of a similarity matrix for the data, instead with DP-means one takes all eigenvectors where the eigenvalues are above a given threshold, then cluster via DP-means using that same threshold to determine when to add additional components. Similarly, the author also shows that DP-means can be extended to graph cut problems. The paper concludes with a few simulations demonstrating the effectiveness of DP-means and its multiple data set extension. \par
This article offers obvious extensions of how the classical Bayesian statistics we learned can be extended to a machine learning algorithm. The prior Dirichlet process is used to guide the posterior number of components. The method for fitting the model is the limit of a Gibbs sampling algorithm. Finally, the multiple data set version of DP-means is a hierarchical model, based on the exchangeability of the data sets. All together, this forms an interesting competitor to k-means, being similarly easy to compute, but built on classic Bayesian principles.

\section{Paper Contents}
\subsection{DP-Means Algorithm}
The idea of clustering arises, in statistical terms, from the idea that a set of random variables are drawn from a mixture distribution. In other words, a random variable $X_i$ is drawn in a two stage process. First, a multinomial variable $z_i$ with $k$ possible outcomes is drawn. Then, $X_i\st z_i\sim F_{z_i}$, where \(\left\{F_j\st j\in\{0,...,k\}\right\}\) is some set of probability distributions. The goal of a clustering method is then to impute the $z_i$ based on the $X_i$, revealing important underlying structure in the data. Bayesian models provide a natural way to fit probability distributions to $z_i$, in particular without choosing $k$ a priori, but there issue is that they, classically, are complicated to compute on large data sets and don't necessarily scale well. \par
Thus, the k-means algorithm, where $k$ must be specified beforehand, remains the most commonly used algorithm. This method can be thought of as designed for a model where $z_i$ is multinomial $k$, for fixed $k$, and then $X\st z_i \sim N_p(\mu_j,\sigma I_p)$ for some set of mean vectors \(\left\{\mu_j\st j\in\{0,...,k\}\right\}\). To find $\{\mu_i\}$ using k-means, one picks initial guesses for the $\hat\mu_j$, then alternates assigning \(\hat z_i = \argmin_j \|X_i - \hat\mu_j\|\) and 
\[\hat\mu_j = \frac{1}{\big | \{i\st z_i=j\} \big |}\sum_{i\st z_i=j }X_i \] 
until the $\mu_j$ and $z_i$ have converged. This method allows little flexibility, but is relatively easy to compute, even for large data sets. The authors of the paper notes that the fitting process is the limit of what the EM algorithm would do if $\sigma$ goes to 0. \par

In designing the DP-means algorithm, the authors ``attempt to achieve the best of both worlds by designing scalable hard clustering algorithms from a Bayesian nonparametric viewpoint." To this end, they start with a purely Bayesian model for how the data would arise which puts a prior on $k$, and then similarly derives a simple algorithm for hard clustering that represents the limit of fitting this Bayesian algorithm as a variance parameter is sent to 0.

>>>>

\newcommand{\p}{\mathrm{P}}
\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\trace}{\mathrm{trace}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\ST}{mbox{ s.t. }}

@@

\newcommand{\st}{ \; \big | \:}
\newcommand{\car}[1]{ \left\vert #1 \right\vert}

@@

\begin{document}
\title{Literature Report on DP-means}
\author{Aaron Maurer}
\date{STAT 302, Spring Quarter 2015}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
For this report, I chose to read ``Revisiting k-means: New Algorithms via Bayesian Nonparametrics"\footnote{Brian Kulis and Michael I. Jordan, ``Revisiting k-means: New Algorithms via Bayesian Nonparametrics'', \textit{CoRR} (2011): http://arxiv.org/abs/1111.0352}. This article introduces an alternative to the k-means clustering algorithm called DP-means, where instead of specifying a number of clusters at the start, then working to optimize their fit, new clusters are introduced when a point in the data set is too far from one of the existing clusters. This in itself is a `hard' clustering algorithm, which outputs an assignment to clusters rather than a posterior distribution across clusters typical of a Bayesian model. However, this paper demonstrates that this algorithm is the limit of a Bayesian model. In the model, the total distribution is a mixture of Gaussian distributions, with the number of components coming from a Dirichlet process. This is the source of the name of the algorithm: Dirichlet process means or DP-means. \par
In addition to a description and derivation of this algorithm, the paper offers a few extensions and simulations. The most interesting of theses is an extension of DP-means called the ``hard Gaussian hierarchical Dirichlet process", over multiple data sets, local clusters are simultaneously fit on each data set to match a set of global clusters. This is again a hard clustering algorithm, but it can be derived as well in a Bayesian fashion by taking the limit of a Bayesian model. The next extension shows how DP-means extends to spectral clustering in a similar fashion to how k-means does. Similarly, the author also shows that DP-means can be extended to graph cut problems. The paper concludes with a few simulations demonstrating the effectiveness of DP-means and its multiple data set extension. \par
This article offers obvious extensions of how the classical Bayesian statistics we learned can be extended to a machine learning algorithm. The DP-means algorithm is a hierarchical model which uses a Dirichlet process as the prior on the number of clusters. The method for fitting the model is the limit of a Gibbs sampling algorithm. The hard Gaussian HDP extension is a typical Bayesian extension of the original model, where an additional level prior distributions is added to an existing model. All together, this forms an interesting competitor to k-means, being similarly easy to compute, but built on classic Bayesian principles.

\section{Paper Contents}
\subsection{DP-Means Algorithm}
The idea of clustering arises, in statistical terms, from the idea that a set of random variables are drawn from a mixture distribution. In other words, a random variable $X_i$ is drawn in a two stage process. First, a multinomial variable $z_i$ with $k$ possible outcomes is drawn. Then, $X_i\st z_i\sim F_{z_i}$, where \(\left\{F_j\st j\in\{0,...,k\}\right\}\) is some set of probability distributions. The goal of a clustering method is then to impute the $z_i$ based on the $X_i$, revealing important underlying structure in the data. Bayesian models provide a natural way to fit probability distributions to $z_i$, in particular without choosing $k$ a priori, but there issue is that they, classically, are complicated to compute on large data sets and don't necessarily scale well. \par
Thus, the k-means algorithm, where $k$ must be specified beforehand, remains the most commonly used algorithm. This method can be thought of as designed for a model where $z_i$ is multinomial $k$, for fixed $k$, and then $X_i\st z_i \sim N_p(\mu_c,\sigma I_p)$ for some set of mean vectors \(\left\{\mu_c\st c\in\{0,...,k\}\right\}\). To find $\{\mu_i\}$ using k-means, one picks initial guesses for the $\hat\mu_c$, then alternates assigning \(\hat z_i = \argmin_c \|X_i - \hat\mu_c\|\) and 
\[\hat\mu_c = \bar X_c = \frac{1}{\car{\{i\st z_i=c\}}}\sum_{i\st z_i=c }X_i \] 
until the $\mu_c$ and $z_i$ have converged. This method allows little flexibility, but is relatively easy to compute, even for large data sets. The authors of the paper notes that the fitting process is the limit of what the EM algorithm would do for a mixture of $k$ Gaussians if the variance of each, $\sigma$, goes to 0. \par

In designing the DP-means algorithm, the authors ``attempt to achieve the best of both worlds by designing scalable hard clustering algorithms from a Bayesian nonparametric viewpoint." To this end, they start with a purely Bayesian model for how the data would arise which puts a prior on $k$, and then similarly derives a simple algorithm for hard clustering that represents the limit of fitting this Bayesian algorithm as a variance parameter is sent to 0. The Bayesian model for how the data arises sequentially is something like this: 
\begin{itemize}
    \item The $z_i$ arise from a Dirichlet process. This means that given a parameter $\alpha$, for the $i$th draw, if we have $k$ clusters already
        \begin{itemize}
            \item $z_i=c$ for $c\leq k$ with probability \(\frac{\car{\{ j\vert j<i, z_j=c\}}}{\alpha + i -1}\)
            \item $z_i=k+1$ with probability \(\frac{\alpha}{\alpha + i -1}\)
        \end{itemize}
    \item The cluster centers $\mu_1,...,\mu_k\sim N(0,\rho I)$ for some variance parameter $\rho$.
    \item Finally, \(X_i\st z_i \sim N(\mu_{z_i},\sigma I)\) for another variance parameter $\sigma$.
\end{itemize}
The authors fix $\rho$ and $\sigma$, and then set $\alpha=\left(1+\frac{\rho}{\sigma}\right)^{d/2}\exp\left(-\frac{\lambda}{2\sigma}\right)$, where $d$ is the dimension of the $X_i$. To set up a Gibbs sampling algorithm, we can calculate the conditional distributions of the unknown parameters $z_1,...,z_n$ and $\mu_1,...,\mu_k$. Let us say that there are $k$ clusters, excluding $z_i,X_i$, at any given point. Also, let the number of points assigned to cluster $c$ excluding $z_i,X_i$ is $n_{-i,c}$. This yields, for the conditional distribution of the $z_i$,
\begin{align*} 
    \p(z_i=c \st X_{1:n},z_{-i},\mu_{1:k},\rho,\sigma) &= \p(z_i=c \st X_i,z_{-i},\mu_{1:k},\rho,\sigma) \\
                                           &\propto \begin{cases} 
                                                \p(X_i\st z_i=c,\sigma,\rho)\p(z_i=c\st z_{-i})& \mbox{if } c=k+1 \\ 
                                                \p(X_i\st z_i=c,\mu_c,\sigma)\p(z_i=c\st z_{-i})& \mbox{if } c\leq k \\
                                            \end{cases} \\
                                           &\propto \begin{cases} 
                                                (2\pi\sigma)^{-d/2} \exp\left(-\frac{\lambda}{2\sigma}-\frac{\|X_i\|^2}{2(\rho+\sigma)}\right) & \mbox{if } c=k+1 \\
                                                n_{-i,c}(2\pi\sigma)^{-d/2} \exp\left(-\frac{\|X_i-\mu_c\|^2}{2\sigma}\right) & \mbox{if } c\leq k \\
                                            \end{cases}\\
                                           &\propto \begin{cases} 
                                                \exp\left(-\frac{\lambda}{2\sigma}-\frac{\|X_i\|^2}{2(\rho+\sigma)}\right) & \mbox{if } c=k+1 \\
                                                n_{-i,c}\exp\left(-\frac{\|X_i-\mu_c\|^2}{2\sigma}\right) & \mbox{if } c\leq k \\
                                            \end{cases}
\end{align*}
For the conditional distribution of the $\mu_c$ given everything else, it yields, where $\bar X_c$ is the mean over all $X_i\st z_i=c$,
\begin{align*}
    \p(\mu_c=\tau\st X_{1:n},z_{1:n},\mu_{-i},\rho,\sigma) &\propto \p( \bar X_c \st n_c,\mu_c, \sigma)\p(\mu_c=\tau\st \rho) \\
    &\propto \exp\left(\frac{1}{(\rho^{-2}+n\sigma^{-2})}\left\|\bar X_c - \frac{n_c\bar X_c}{\sigma^2(\rho^{-2}+n\sigma^{-2})}\right\|^2\right)
\end{align*}
Now, as one sends $\sigma$ to $0$, both of these probabilities become point masses. For the first probability, $\p(z_i=c \st X_{1:n},z_{-i},\mu_{1:k},\rho,\sigma)$, it is a point mass on $z_i=c\leq k$ if $\|\mu_c-X_i\|$ is minimal and less than $\lambda$, otherwise it is a point mass on $z_i=k+1$. The second distribution, of $ \mu_c\st X_{1:n},z_{1:n},\mu_{-i},\rho,\sigma$, is just a point mass on $\bar X_c$. Thus, the Gibbs sampling algorithm initialized with one cluster at the global mean turns into the DP-means algorithm, which is iterated till convergence:
\begin{itemize}
    \item Initialize $k=1$, $\mu_1 = \bar X$, and $z_i=1 \;\forall i$
    \item For each $X_i$
        \begin{itemize}
            \item {\bf If $X_i$ is too far from a cluster, introduce a new one:} if \(\lambda \leq \|X_i-\mu_c\|^2 \; \forall c\), set $k=k+1$, $z_i=k$ and $\mu_k=X_i$
            \item {\bf Otherwise, have it join the nearest cluster:} Otherwise $z_i=\argmin_c\|X_i-\mu_c\|^2$
        \end{itemize}
    \item {\bf Set the cluster centers to the mean of the cluster:} For each $c$, set $\mu_c = \bar X_c$
\end{itemize}
\subsection{Hierarchical DP-means}
As is common with Bayesian models, a hierarchical model such as DP-means can be expanded to a more complicated model by adding an additional level of hierarchy by putting a probability distribution on some of the hyperparameters. This is exactly what the authors of this paper do to DP-means. They replace the original fixed Dirichlet process with a mixture of Dirichlet processes as determined by an additional Dirichlet process. This is called a hierarchical Dirichlet process. While this doesn't yield anything interesting for a single data set, this allows for a clustering algorithm that clustering over multiple data sets. The idea is that one might have multiple data sets of different observations, but with identical measurements taken for each, where each data set has a set of local clusters within it that don't necessarily align with the local clusters of another data set. This method will allow the simultaneous identification of a set of global clusters while the data sets borrow strength from each other to identify the local clusters within each data set. \par
The ``Revisiting k-means:..." paper doesn't go into the full derivation of this model from a Bayesian setup, instead pointing to the original paper on the hierarchical Dirichlet process\footnote{Yee Teh, Micheal Jordan, Matthew Beal, David Blei, ``Hierarchical Dirichlet Processes", JASA, 1001(476):1566-1581, 2006.} and indicating that ``the limiting process described earlier for the standard DP can be straightforwardly extended to the HDP". To start, the Bayesian model for how the data arises goes something like this:
\begin{itemize}
    \item The $z_{ij}$ which are the local indicators of the group for the $i$th observation in the $j$th data set, are drawn first. Lets say, up to the current $z_{ij}$ being drawn, there are $k_{j.}$ groups in the data set, each group has $n_{jc}$ observations in it, and $z_{ij}$ is the $n_{j.}$th observation drawn. $z_{ij}$ is drawn locally by a Dirichlet process with parameter $\alpha$. 
        \begin{itemize}
            \item As before, $z_{ij}=c\leq k_{j.}$ with probability $\frac{n_{jc}}{\alpha+n_{j.}-1}$
            \item $z_{ij}=k_j+1$ with probability $\frac{\alpha}{\alpha+n_{j.}-1}$
        \end{itemize}

    \item $v_{jc}$, which is the association of the $c$th group in the $j$th data set with a global cluster, is drawn next. Let us say that there are $g$ global clusters, and $k_{.p}$ is the number of groups accross all  data set $j$ for which $v_{jc}=p$. $v_{jc}$ is drawn globally by a second Dirichlet process with parameter $\gamma$:
        \begin{itemize}
            \item $v_{jc}=p\leq g$ with probability $\frac{k_{.p}}{\gamma+k_{..}-1}$
            \item $v_{jc}=g+1$ with probability $\frac{\gamma}{\gamma+k_{..}-1}$
        \end{itemize}
    \item All the global cluster centers $\mu_p$ are drawn iid from $N(0,\rho I)$ for some variance parameter $\rho$
    \item Finally, each observed data point $X_{ij}\st z_{ij},v_{jz_{ij}} \sim N(\mu_{v_{jz_{ij}}}, \sigma I)$ for another variance parameter $\sigma$
\end{itemize}
After fixing $\sigma$ and $\rho$ in a similar manner to before, the conditionals can be constructed for a Gibbs sampler. Then $\alpha$ and $\gamma$ are in terms of local and global parameter $\lambda_l$ and $\lambda_g$. In the same way, as $\sigma$ is sent to $0$, this makes each conditional distribution becomes a point mass. This turns the Gibbs sampler into the "Hard Gaussian HDP", which is the hard clustering algorithm to fit the local and global clusters across data sets. The algorithm is this, iterated till convergence.
\begin{itemize}
    \item Initialize $g=1$, $k_j=1$, $\mu_1$ = $\bar X$, $z_{ij}=1$, $v_{j1}=1$ for all $i$ and $j$.
    \item For each $X_{ij}$:
        \begin{itemize}
            \item {\bf If the point is too far from a global cluster, introduce a new one:}  \\
                If
                    \[\lambda_l<\|X_{ij}-\mu_{v_{jc}}\|^2 \;\forall c<k_j \mbox{ and } \lambda_l+\lambda_p<\|X_{ij}-\mu_{p}\|^2 \;\forall p\leq g\]
                    set $k_j=k_j+1$, $g=g+1$, $z_{ij}=k_j$, $v_{jk_j}=g$, $\mu_g=X_{ij}$
            \item {\bf If the point is too far away from a local cluster, but near a global cluster, add a new local cluster:} \\
                If 
                \[\min_c\|X_{ij}-\mu_{v_{jc}}\|^2>\min_p\|X_{ij}-\mu_{p}\|^2+\lambda_l\]
                set \(k_j=k_j+1$, $z_{ij}=k_j$, and $v_{jz_{ij}}=\argmin_p\|X_{ij}-\mu_{p}\|^2$
            \item {\bf Otherwise, match to one of the nearest clusters:}  \\
                Set $z_{ij}=\argmin_c\|X_{ij}-\mu_{v_{jc}}\|^2$
        \end{itemize}
    \item For each local cluster $c_j$:
        \begin{itemize}
            \item{\bf If the local cluster is too far from a global cluster, introduce a new one:}
                If 
                \[\min_p\sum_{\{X_{ij}\st z_{ij}=c_j\}}\hspace{-1em} \|x-\mu_p\|^2\;>\;\lambda_l+ \hspace{-1em} \sum_{\{X_{ij}\st z_{ij}=c_j\}}\hspace{-1em}\|x-\bar X_{jc_j}\| \]
                Set $g=g+1$, $v_{jc_j}=g$, and $\mu_g=\bar X_{jc_j}$
            \item{\bf Otherwise, match each the local cluster to the nearest global cluster:}
                Set 
                \[v_{jc_j} = \argmin_p\sum_{\{X_{ij}\st z_{ij}=c_j\}}\hspace{-1em} \|x-\mu_p\|^2\]
        \end{itemize}
    \item For each global cluster $p$:
        \begin{itemize}
            \item {\bf Recenter each global cluster:}\\
                Set
                \[ \mu_p = \sum_{\{X_{ij}\st v_{jz_{ij}}=p\}}\hspace{-1em} x \]
        \end{itemize}
\end{itemize}
Its worth noting that the final $z_{ij}$ aren't meaningful in the end, since observations are allocated to local clusters only based on the associated global cluster center. This means that the partition of points into local clusters which map to the same global cluster is nonidentifiable. Thus, the important value is the global cluster association of each point, $v_{jz_{ij}}$.
\subsection{Spectral Clustering with DP-means}
A convieient way of thinking of clustering is that one is trying to find the clustering such that when one replaces the observations with the clusters, the result is the minimal distortion of the data. Another way of saying this is that we want to project our data onto a set of vectors in the fashion that reduces the sum of the norms, or sum of squares, the least. This is akin to linear regression, where we are looking to project our data onto a lower dimmensional plane in a similar fashion. Just as with linear regression, where this projection was performed with the hat matrix $H$, for each clustering we can use a weight matrix $W$ to do this. $W_{ij}=\frac{1}{n_c}$ if $X_i$ and $X_j$ are to be placed in the same cluster, and $0$ otherwise. This means $WX$ replaces each row of $X$, so we can calculate are total distortion as $\trace(X'X)-\trace(W'X'XW)$. Since all the rows in $WX$ will be identical within a given cluster, we can replace this by a weighted sum of the unique cluster centers. Thus, $\trace(W'X'XW)=\trace(Y'X'XY)$, where $Y$ is the normalized $n\times k$ clustering matrix, which will have $Y_{ic}=\frac{1}{\sqrt n_c}$ if observation $i$ is in cluster $c$, and 0 otherwise. \par
Both clustering algorithms, k-means and DP-means, attempt to minimize this distortion, or maximize $\trace(Y'X'XY)$, under a certain criteria. For k-means, that criteria is over all normalized clustering matricies $Y$ with $k$ is fixed, while for DP-means, it turns out the criteria is to minimize $\trace(Y'X'XY)-\lambda k$ over all normalized clustering matricies $Y$ and corresponding $k$. It turns out that a natural relaxation, or simplification of both of these targets is to let $Y$ be any orthonormal matrix rather than just a normalized clustering matrix. This is where the spectral part comes in; let $X'X=Q\Lambda Q'$ be the eigen decomposition of $X$, where $Q$ is the orthonormal matrix of eigenvalues and $\Lambda=\diag(\lambda_1,...,\lambda_n)$ is the diagonal matrix of eigenvalues in decreasing order. Its obvious that the optimal $Y$ in both cases will be a subset of the columns of $Q$. For k-means, it will be the the first $k$ columns, corresponding to the largest $k$ eignvalues, while for DP-means, it will be all columns such that $\lambda_p>\lambda$. \par
The reason this relaxation is useful is two fold. First, when one has the relaxed matrix $Y$ above, they can perform a potentially simpler clustering on it, treating it as a new data matrix. This clustering will also work very well on the original data set, since $Y$ is a set of principle components from it. More importantly though, if one thinks of $X'X$ as a matrix meauring similarity between all the observations in a particular fashion, it can be replaced with another similarity matrix $K$ which uses a different critiera, for instance $K_{ij} = \exp\left(-\|X_i-X_j\|^2\right)$. Then, one similarly can make a matrix $Y$ which is the first $k$ eigenvectors for k means or all eigenvectors with eigenvalues greater than $\lambda$ for DP-means. The advantage of this is when $X$ is not a nice mix of gaussian clusters, as needed for k-means or DP-means to work well, one can use a different measure of similarity to generate $K$ and $Y$ such that $Y$ is approximately a nice mixture of Gaussians.

\subsection{Remaining Content}
The remaining part of the paper had two sections. The first of these has to do with graph clustering. The paper only gives a very brief desciription of the result, without enough detail to understand any of it without knowing the area already. The gist is that there is a natural extension of DP-means to the graph cut problem, just as there is with k-means. It turns out that, as has been demonstrated similarlily in the three earlier sections, DP-means ends up searching for an equivalent optimal result as k-means given the right parameterization. \par
The final section deals with a number of simulations that the authors conducted comparing fitting models by k-means, DP-means, and a full Dirichlet process mixture with Gibbs sampling. On individual data sets, they all generally produce similar quality clusters. On large data sets though, Gibbs sampling becomes computationally infeasible while both k-means and DP-means run in similar time. The impressive result though come from the hard Gaussian HDP. To test it, they created 15 cluster centers, then created 50 data sets which each had 5 data points generated from 5 randomly chosen clusters. The hard Gaussian HDP clustering, fit accross all data sets, performed considerably better than k-means or DP-means fit on all the data sets together or each data set individually.

\section{Similarities to Class Material}
The central idea which this paper uses that we learned in class is that of the hiearchical model. In the based DP-algorithm, the underlying Bayesian model treats the observations $X_i$ as exchangable, allowing the $X_i$, although independent conditioned on their clusters, to lend strength to each other in infering both their cluster associations and the cluster means. This concept, where the observations share information, is a central one we saw in many contexts in class. The hard Gaussian HDP model displayed another precept of heiarchical models we saw inclass; Bayesian approaches are accomodating to additional levels of hiearchy being added to models. This is exactly how the hard Gaussian HDP was developed. They added an additional level of hiearchy to the DP-means model so that observations wouldn't just lend strength to each other within a data set but also accross data sets.\par
The second important idea we saw before in class was fitting the model by a MCMC process, in particular Gibbs sampling. There model was complicated to the point that deriving an analytical solution would be at best infeasible. Since the marginal distributions of parameters could be expressed in closed form, this allowed a Gibbs sampler to be set up. An interesting twist I thought was stretching the Gibbs Sampler till it was deterministic so as to derive a hard clustering algorithm. While it wasn't like the other markov chains we saw which were irreducible, instead having an absorbing state, it still proved useful in fitting the model. \par
The final idea from class that stuck out was the idea of an uniformative prior. While this isn't how it was described in the paper, several of the priors struck me as being set up to be largely uninformative. For instance, by sending the within cluster variance to 0 while the cluster mean variance was fixed, the result was essentially a uniform prior on $\R^p$ for the cluster means, such that within a cluster the center was just set to the mean. Certainly, taking the limit of a distribution as a parmeter goes to 0 or infinity was a common trick we saw to set up uninformative priors.
\section{My Thoughts}
On the whole, I found this paper very interesting to read, and learned a lot from it and the related papers I looked up. However, as a practical matter, they established DP-means as a largely equivalent method to k-means rather than a superior method. Throught the paper, they showed that they the two methods acheived largely similar results in similar times in several different areas where they could both be used. The main difference was in how they are parameterized. In practice, with either one, I would want to try several different values of the parameter and compare the fit. I'm not sure why DP-means would make deciding on the proper parameter easier. I would just be looking for a number of parameters to map to the same number of clusters with DP-means, while with k-means I would be looking to a falloff in the marginal improvement in distortion. I like that its parameter, $\lambda$ is akin to the complexity parameter we see in a lot of machine learning algorithms, but that is more from an aesthetic point of view than anything else. \par
The value of the Bayesian approach, for me, comes with the hard Gaussian HDP algorithm. As mentioned before, the adding hiearchy to Bayesian models is natural to the extent it often isn't with other models, which is why there is not, or I am unaware of, a k-means based algorithm which can similarly be fit simaltaneously accross many data sets. I would certainly choose this algorithm if I ran into a situation where I was trying to fit clusters accross many data sets and had reason to expect that not all clusters were within all data sets. That said, I think that situation is a somewhat rare occurance. Either way, it is a good tool to have, and learning about hiearchical Dirichlet processes through this algorithm was interesting. \par
An extension which I would have liked to see in the paper would be an algorithm for hiearchical clustering in the more traditional sense, where the base clusters within one data set are clustered into meta clusters within that one data set. For instance, one might imagine cluster centers $\mu_{ij}\sim N(\nu_i,\sigma I)$ with $\nu_i\sim N(0,\rho I)$. This is fertile grounds for a similar Bayesian algorithm based on $DP-means$, and it seems like something which wouldn't be hard to develop.
\section{Big Picture}
This is a fairly recent paper, adding to the relatively recent area of Bayesian nonparametrics which I know at least one of the authors, Micheal Jordan, has been fairly central to. The uniting thread of this field seems to be attempts to model objects which can have infinite dimmension with a countably infinite, nonparametric discrete distrubtion. The dirichlet process, which is the conjugate prior to this family of distributions, thus tends to pop up a lot, as it does here. This process, though a newer development, is based on the Dirichlet distribution, which has long been a tool of traditional Bayesian statistics. Bayesian nonparametrics is also closely tied to the more general statistics/computer science field of machine learning, with many of the methods of the former being adopted by the latter. This paper was, in large part, an attempt of the authors to push this process further by offering a bayesian nonparametric algorithm which has appealing properties to machine learning, mainly scalability.
